```{r setup}
#| include: false
library(tidyverse)
```

# Benchmark {.unnumbered}

We quickly reported benchmarks of the tool for other users. Most rules takes less than a minute except the one retrieving data. CORDEX and CMIP6 retrieval takes around 7 minutes and almost always less than 10 minutes. CHELSA 2 retrieval takes more than 20 minutes but is only done once. Similarly most rules are using less than 1 GB of memory, except CMIP6 retrieval that approach 10 GB, CORDEX retrieval above 20 GB, and CHELSA 2 retrieval above 30 GB. Finally most rules write less than 1 GB, CORDEX and CMIP6 around 5 GB, and CHELSA more than 20 GB, but a lot are only temporary files.

```{r prep}
#| eval: false
bc <- lapply(list.files("results/benchmarks/", full.names = T), readr::read_tsv) %>% 
  bind_rows() %>% 
  mutate(file = list.files("results/benchmarks/"))
bc <- bc %>% 
  separate(file, c("rule1", "rule2")) %>% 
  mutate(rule = paste(rule1, rule2)) %>% 
  mutate(rule = recode(rule,
                       "downscale bc" = "downscale",
                       "eval C" = "eval",      
                       "eval French" = "eval",        
                       "eval New" = "eval",      
                       "eval Vanuatu" = "eval",      
                       "get area" = "area",    
                       "get chelsa2" = "CHELSA",
                       "get chirps" = "CHIRPS",
                       "get gshtd" = "GSHTD",
                       "get cmip6" = "CMIP6",   
                       "get cordex" = "CORDEX", 
                       "hist base" = "hist",
                       "hist ds" = "hist",
                       "merge eval" = "merge",
                       "merge hist" = "merge",
                       "tmf C" = "tmf",
                       "tmf French" = "tmf",
                       "tmf New" = "tmf"))
write_tsv(bc, file = "outputs/bencmark.tsv")
```

```{r time}
#| message: false
#| warning: false
#| fig-cap: "Rules execution time."
read_tsv("outputs/bencmark.tsv") %>% 
  ggplot(aes(s/60, rule, label = `h:m:s`)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  scale_x_sqrt() +
  xlab("Minutes") + ylab("")
```

```{r mem}
#| message: false
#| warning: false
#| fig-cap: "Rules Maximum “Virtual Memory Size”, this is the total amount of virtual memory used by the process."
read_tsv("outputs/bencmark.tsv") %>% 
  ggplot(aes(max_vms/10^3, rule)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  scale_x_sqrt() +
  xlab("Maximum Virtual Memory Size (GB)") + ylab("")
```

```{r io_out}
#| message: false
#| warning: false
#| fig-cap: "Rules writting volum."
read_tsv("outputs/bencmark.tsv") %>% 
  ggplot(aes(io_out/10^3, rule)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  scale_x_sqrt() +
  xlab("GB Written") + ylab("")
```
