```{r set}
#| include: false
library(vroom)
library(tidyverse)
library(terra)
```

# Tropical Moist Forests {.unnumbered}

Tropical moist forest (TMF) climatic niche distributions were projected at the end of the century (2070-2100) using downscaled CMIP6 and CORDEX projections for scenario 2.6 and 8.5. Specifically, we used the definition from Holdridge 1967 [@holdridge1967] as areas with a mean annual temperature over 16°C, a total annual precipitation over 1,000 $mm~yr^{-1}$, and a ratio of potential evapotranspiration over precipitation below 1. Potential evaportanspiration (pet, $mm~month^{-1}$) was derived from monthly mean, maximum and minimum temperatures (tas, tasmax, tasmin, °C) and surface downward short-wave radiation (rsds, $MJ~m^{-~2}~month^{-1}$) as:

$$
 pet = 0.0023 \times rsds \times (tas + 17.8) \times (tasmax - tasmin)^{0.5}
$$

Surface downward short-wave radiation was derived from CHELSA 2 historical climatologies and assumed constant.

## Current

```{r current_tmf_data}
#| eval: false
files <- list.files("outputs/tmf_hist", full.names = TRUE)
names(files) <- c("Côte-d'Ivoire", "French-Guiana", "New-Caledonia")
tmf <- files %>%
  lapply(rast) %>%
  lapply(as.data.frame, xy = TRUE) %>%
  bind_rows(.id = "area") %>%
  mutate(y = ifelse(area == "New-Caledonia", -y, y)) %>%
  na.omit()
write_tsv(tmf, "outputs/current_tmf.tsv")
```

```{r current_tmf_map}
#| message: false
#| warning: false
#| fig-cap: "Fig. S7. Current distribution of the climate niche of the tropical moist forests in Ivory Coast, French Guiana and New Caledonia. The climate niche has been defined based on CHELSA2 data for the 1980-2005 period using Holdridge (1967)’s definition of areas with a mean annual temperature over 16°C, a total annual precipitation over 1,000 mm, and a ratio of potential evapotranspiration over precipitation below 1. Potential evapotranspiration (pet) was derived from monthly mean, maximum and minimum temperatures (tas, tasmax, tasmin) and surface downward short-wave radiation (rsds) as pet = 0.0023 x rsds x (tas+17.8) x (tasmax - tasmin)^0.5." #nolint
g <- read_tsv("outputs/current_tmf.tsv") %>%
  mutate(area = ifelse(area == "Côte-d'Ivoire", "Ivory-Coast", area)) %>%
  mutate(area = gsub("-", " ", area)) %>%
  ggplot(aes(x, y, fill = as.character(tmf))) +
  geom_raster() +
  facet_wrap(~area, scales = "free") +
  theme_bw() +
  scale_fill_manual(guide = "none", values = c("darkgrey", "#48712d")) +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.spacing = unit(0, "lines"),
    aspect.ratio = 1
  )
ggsave("figures/fs07.png", g, dpi = 800, bg = "white", width = 8)
g
```

## Projected

```{r preptmf}
#| eval: false
index <- data.frame(
  file = list.files("results/tmf/nc/", full.names = FALSE),
  path = list.files("results/tmf/nc/", full.names = TRUE)
) %>%
  separate(file, c(
    "country", "project", "domain", "institute", "model", "experiment",
    "ensemble", "rcm", "downscaling", "base", "aggregation",
    "period_future", "period_present", "period_hist"
  ), sep = "_", remove = FALSE)
files <- index %>%
  select(country, project, experiment) %>%
  unique()
for (r in seq_len(nrow(files))) {
  filter(
    index,
    country == files$country[r],
    project == files$project[r],
    experiment == files$experiment[r]
  )$path %>%
    lapply(rast, lyrs = "tmf") %>%
    rast() %>%
    sum() %>%
    writeRaster(
      filename = paste0(
        "outputs/tmf/",
        files$country[r], "_",
        files$project[r], "_",
        files$experiment[r], ".tif"
      ),
      overwrite = TRUE
    )
}
```

```{r tmf_data}
#| eval: false
fg <- rast(c(
  list.files("outputs/tmf/", pattern = "Guiana", full.names = TRUE),
  "outputs/tmf_hist/French-Guiana.nc"
))
names(fg) <- c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5", "hist")
fg_tab <- as.data.frame(fg, xy = TRUE, na.rm = FALE) %>%
  mutate_at(
    c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5"),
    ~ . / max(abs(.), na.rm = TRUE)
  ) %>%
  gather(projection, tmf, -x, -y, -hist) %>%
  separate(projection, c("project", "experiment"), "_") %>%
  mutate(project = factor(project, levels = c("CMIP6", "CORDEX"))) %>%
  na.omit() %>%
  mutate(agreement = abs(tmf) * 100) %>%
  mutate(type = ifelse(tmf >= 0, "gain", "loss")) %>%
  mutate(type = ifelse(tmf == 0, "stable", type)) %>%
  mutate(type = ifelse(hist == 0 & tmf == 0, "non-forest", type)) %>%
  mutate(agreement = ifelse(type %in% c("stable", "non-forest"),
    100, agreement
  ))
write_tsv(fg_tab, "outputs/fg_tmf.tsv")
ci <- rast(c(
  list.files("outputs/tmf/", pattern = "Ivoire", full.names = TRUE),
  "outputs/tmf_hist/Côte-d'Ivoire.nc"
))
names(ci) <- c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5", "hist")
ci_tab <- as.data.frame(ci, xy = TRUE, na.rm = FALSE) %>%
  mutate_at(
    c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5"),
    ~ . / max(abs(.), na.rm = TRUE)
  ) %>%
  gather(projection, tmf, -x, -y, -hist) %>%
  separate(projection, c("project", "experiment"), "_") %>%
  mutate(project = factor(project, levels = c("CMIP6", "CORDEX"))) %>%
  na.omit() %>%
  mutate(agreement = abs(tmf) * 100) %>%
  mutate(type = ifelse(tmf >= 0, "gain", "loss")) %>%
  mutate(type = ifelse(tmf == 0, "stable", type)) %>%
  mutate(type = ifelse(hist == 0 & tmf == 0, "non-forest", type)) %>%
  mutate(agreement = ifelse(type %in% c("stable", "non-forest"),
    100, agreement
  ))
write_tsv(ci_tab, "outputs/ci_tmf.tsv")
nc <- rast(c(
  list.files("outputs/tmf/", pattern = "Caledonia", full.names = TRUE),
  "outputs/tmf_hist/New-Caledonia.nc"
))
names(nc) <- c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5", "hist")
nc_tab <- as.data.frame(nc, xy = TRUE, na.rm = FALSE) %>%
  mutate_at(
    c("CMIP6_2.6", "CMIP6_8.5", "CORDEX_2.6", "CORDEX_8.5"),
    ~ . / max(abs(.), na.rm = TRUE)
  ) %>%
  gather(projection, tmf, -x, -y, -hist) %>%
  separate(projection, c("project", "experiment"), "_") %>%
  mutate(project = factor(project, levels = c("CMIP6", "CORDEX"))) %>%
  na.omit() %>%
  mutate(y = -y) %>%
  mutate(agreement = abs(tmf) * 100) %>%
  mutate(type = ifelse(tmf >= 0, "gain", "loss")) %>%
  mutate(type = ifelse(tmf == 0, "stable", type)) %>%
  mutate(type = ifelse(hist == 0 & tmf == 0, "non-forest", type)) %>%
  mutate(agreement = ifelse(type %in% c("stable", "non-forest"),
    100, agreement
  ))
write_tsv(nc_tab, "outputs/nc_tmf.tsv")
```

<!-- Surprisingly, CORDEX projections systematically predicted reduced TMF climatic distribution for both experiment 2.6 and 8.5 compared to CMIP6, with more forest near the coast and in the west following the east to west rainfall gradient. However, in both CORDEX and CMIP6 and both experiments at least one projection was always predicting TMF in every cell. The difference between the two experiments was also a reduced TMF distribution under higher temperature of 8.5, with more forest near the coast and in the west following the east to west rainfall gradient. -->

```{r tmffg}
#| message: false
#| warning: false
#| fig-cap: "Fig. S8. Tropical moist forest ditribution predicted in French Guiana by the end of the century for CMIP6 and CORDEX model with scenario 2.6 and 8.5." #nolint
g <- vroom("outputs/fg_tmf.tsv") %>%
  mutate(experiment = paste0(experiment, "~W~m^{-~2}")) %>%
  mutate(type = recode(type,
    "stable" = "stable forest",
    "non-forest" = "stable non-forest"
  )) %>%
  ggplot(aes(x, y, fill = type, alpha = agreement)) +
  geom_raster() +
  coord_equal() +
  facet_grid(experiment ~ project, labeller = label_parsed) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(), panel.grid = element_blank(),
    panel.background = element_rect(fill = "white")
  ) +
  scale_fill_manual("Forest cover change:",
    values = c("#e40400", "black")
  ) +
  scale_alpha_continuous("Model Agreement for gain and loss (%)",
                         range = c(.1, 1)) +
  theme(legend.position = "bottom", legend.box = "vertical")
ggsave("figures/fs08.png", g, dpi = 800, bg = "white", width = 8)
g
```

<!-- CORDEX simulations predicted less TMF in the central and south west of Ivory Coast but more near the coast , especially in scenario 2.6. **I need to explore the drivers, determining variables tas, pet, pr to be added. I also need to explore the weird CORDEX patterns in the North.** -->

```{r tmfci}
#| message: false
#| warning: false
#| fig-cap: "Fig. S9. Tropical moist forest ditribution predicted in Ivory Coast by the end of the century for CMIP6 and CORDEX model with scenario 2.6 and 8.5." #nolint
g <- vroom("outputs/ci_tmf.tsv") %>%
  mutate(experiment = paste0(experiment, "~W~m^{-~2}")) %>%
  mutate(type = recode(type,
    "stable" = "stable forest",
    "non-forest" = "stable non-forest"
  )) %>%
  ggplot(aes(x, y, fill = type, alpha = agreement)) +
  geom_raster() +
  coord_equal() +
  facet_grid(experiment ~ project, labeller = label_parsed) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(), panel.grid = element_blank(),
    panel.background = element_rect(fill = "white")
  ) +
  scale_fill_manual("Forest cover change:",
    values = c("#48712d", "#e40400", "black", "darkgrey")
  ) +
  scale_alpha_continuous("Model Agreement for gain and loss (%)",
                         range = c(.1, 1)) +
  theme(legend.position = "bottom", legend.box = "vertical")
ggsave("figures/fs09.png", g, dpi = 800, bg = "white", width = 8)
g
```

<!-- CORDEX and CMIP6 major projected distributions of TMF were in agreement in New Caledonia. Only less models projected TMF in the south east of the Grande Terre in CORDEX while more models projected TMF on the west coast in CORDEX, especially in the central west coast in experiment 8.5. -->

```{r tmfnc}
#| message: false
#| warning: false
#| fig-cap: "Fig. S10. Tropical moist forest ditribution predicted in New Caledonia by the end of the century for CMIP6 and CORDEX model with scenario 2.6 and 8.5." #nolint
g <- vroom("outputs/nc_tmf.tsv") %>%
  mutate(experiment = paste0(experiment, "~W~m^{-~2}")) %>%
  mutate(type = recode(type,
    "stable" = "stable forest",
    "non-forest" = "stable non-forest"
  )) %>%
  ggplot(aes(x, y, fill = type, alpha = agreement)) +
  geom_raster() +
  coord_equal() +
  facet_grid(experiment ~ project, labeller = label_parsed) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(), panel.grid = element_blank(),
    panel.background = element_rect(fill = "white")
  ) +
  scale_fill_manual("Forest cover change:",
    values = c("#48712d", "#e40400", "black", "darkgrey")
  ) +
  scale_alpha_continuous("Model Agreement for gain and loss (%)",
                         range = c(.1, 1)) +
  theme(legend.position = "bottom", legend.box = "vertical")
ggsave("figures/fs10.png", g, dpi = 800, bg = "white", width = 8)
g
```

```{r fig4}
#| message: false
#| warning: false
#| fig-cap: "Fig. 4. Projected climate niche shift of the tropical moist forest by the end of the century. The tropical moist forest climatic niche has been projected for the period 2070-2100 based on downscaled projections from CORDEX and CMIP6 for mean, minimum and maximum monthly temperature and monthly precipitation. The projected climatic niches were compared with the current climatic niche for each model and scenario to calculate shifts with loss (red) or gain (green) of forest. The maps show with colour intensity in Ivory Coast, French Guiana and New Caledonia the agreement between models in shifts for scenario 2.6 with CMIP6 and CORDEX projections. Black pixels indicate areas of stable forest, while grey pixels indicate areas of historical and projection stable non-forest. Current distribution of the climate niche of the tropical moist forests can be found in Supplementary Figure S7, while details on Ivory Coast, French Guiana and New Caledonia for both CORDEX and CMIP6 and scenarios 2.6 and 8.5 can be found in Supplementary Figures S8, S9, and S10.Means across CMIP6 and CORDEX projections for both scenario 2.6 and 8.5 are shown in Fig. S11." #nolint
tmf <- list(
  "French Guiana" = "outputs/fg_tmf.tsv",
  "Ivory Coast" = "outputs/ci_tmf.tsv",
  "New Caledonia" = "outputs/nc_tmf.tsv"
) %>%
  lapply(vroom) %>%
  bind_rows(.id = "country") %>%
  filter(experiment == "2.6")
g <- tmf %>%
  mutate(country = factor(country,
    levels = c(
      "French Guiana",
      "Ivory Coast",
      "New Caledonia"
    )
  )) %>%
  mutate(type = recode(type,
    "stable" = "stable forest",
    "non-forest" = "stable non-forest"
  )) %>%
  mutate(project_country = paste(project, country)) %>%
  ggplot(aes(x, y, fill = type, alpha = agreement)) +
  geom_raster() +
  facet_wrap(~project_country, scales = "free") +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(), panel.grid = element_blank(),
    panel.background = element_rect(fill = "white")
  ) +
  scale_fill_manual("Forest cover change:",
    values = c("#48712d", "#e40400", "black", "darkgrey")
  ) +
  scale_alpha_continuous("Model Agreement for gain and loss (%)",
                         range = c(.1, 1)) +
  theme(legend.position = "bottom", legend.box = "vertical")
ggsave("figures/f04.png", g, dpi = 300, width = 8, height = 7, bg = "white")
g
```

```{r figs11}
#| message: false
#| warning: false
#| fig-cap: "Fig. S11. Projected climate niche shift of the tropical moist forest by the end of the century. The tropical moist forest climatic niche has been projected for the period 2070-2100 based on downscaled projections from CORDEX and CMIP6 for mean, minimum and maximum monthly temperature and monthly precipitation. The projected climatic niches were compared with the current climatic niche for each model and scenario to calculate shifts with loss (red) or gain (green) of forest. The maps show with colour intensity in Ivory Coast, French Guiana and New Caledonia the agreement between models in shifts for scenarios 2.6 and 8.5 with the means across CMIP6 and CORDEX projections. Black pixels indicate areas of stable forest, while grey pixels indicate areas of historical and projection stable non-forest." #nolint
tmf <- list(
  "French Guiana" = "outputs/fg_tmf.tsv",
  "Ivory Coast" = "outputs/ci_tmf.tsv",
  "New Caledonia" = "outputs/nc_tmf.tsv"
) %>%
  lapply(vroom) %>%
  bind_rows(.id = "country") %>%
  select(country, x, y, hist, experiment, project, tmf) %>%
  pivot_wider(names_from = project, values_from = tmf) %>%
  rowwise() %>%
  mutate(tmf = mean(c(CMIP6, CORDEX))) %>%
  ungroup() %>%
  select(-CMIP6, -CORDEX) %>%
  mutate(agreement = abs(tmf) * 100) %>%
  mutate(type = ifelse(tmf >= 0, "gain", "loss")) %>%
  mutate(type = ifelse(tmf == 0, "stable", type)) %>%
  mutate(type = ifelse(hist == 0 & tmf == 0, "non-forest", type)) %>%
  mutate(agreement = ifelse(type %in% c("stable", "non-forest"),
    100, agreement
  ))
g <- tmf %>%
  mutate(type = recode(type,
    "stable" = "stable forest",
    "non-forest" = "stable non-forest"
  )) %>%
  mutate(experiment = recode(as.character(experiment),
    "2.6" = "~2.6~W~m^{-~2}",
    "8.5" = "~8.5~W~m^{-~2}"
  )) %>%
  mutate(country = gsub(" ", "~", country)) %>%
  mutate(experiment_country = paste0(country, experiment)) %>%
  mutate(experiment_country = factor(experiment_country,
    levels = c(
      "French~Guiana~2.6~W~m^{-~2}",
      "Ivory~Coast~2.6~W~m^{-~2}",
      "New~Caledonia~2.6~W~m^{-~2}",
      "French~Guiana~8.5~W~m^{-~2}",
      "Ivory~Coast~8.5~W~m^{-~2}",
      "New~Caledonia~8.5~W~m^{-~2}"
    )
  )) %>%
  ggplot(aes(x, y, fill = type, alpha = agreement)) +
  geom_raster() +
  facet_wrap(~experiment_country, scales = "free", labeller = label_parsed) +
  theme_bw() +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.border = element_blank(), panel.grid = element_blank(),
    panel.background = element_rect(fill = "white")
  ) +
  scale_fill_manual("Forest cover change:",
    values = c("#48712d", "#e40400", "black", "darkgrey")
  ) +
  scale_alpha_continuous("Model Agreement for gain and loss (%)",
                         range = c(.1, 1)) +
  theme(legend.position = "bottom", legend.box = "vertical")
ggsave("figures/fs11.png", g, dpi = 800, bg = "white", width = 8)
g
```
